Lyman-$\alpha$ photons through rotating outflows

### 2018-07-23 paper/paper.tex
abstract

\begin{abstract}
Outflows and rotation are two ubiquitous kinematic features in the gas
kinematics of galaxies.
Here we perform Monte Carlo radiative transfer simulations of outflowing
gas with additional solid body rotation to understand how these kinematic
features impact the morphology of the Lyman-$\alpha$ emission line.
We explore a range of neutral Hydrogen optical depth of
$10^5\leq\tauh\leq10^7$, rotational velocity $0\leq \vrot/\kms \leq 100$ and
outflow velocity $0\leq \vout/\kms\leq 50$.  
We find three important consequences of rotation.
First, it introduces a dependency with viewing angle; second it
produces a line broadening and third it increases the flux at the
line's center.
We also test a semi-analytic model that modifies the spectra of
pure outflow simulations and manages to reproduce the radiative
transfer results for the line widths and flux change at the line's
center within a $7\%$ and $50\%$ precision, respectively, for an
optical depth of $\tauh=10^5$ and within $2\%$ and $1\%$ for an
optical depth of $\tauh=10^7$.   
We argue that rotation is a required kinematic feature to take into
account for a correct interpretation of \lya spectra in systems where
rotation is already evident through kinematic maps but also in other
systems, such as Green Pea galaxies, where a strong \lya emission at
the line's center could be explained by rotational kinematics. 
\end{abstract}

>>>>

\begin{abstract}
Outflows and rotation are two ubiquitous kinematic features in the gas
kinematics of galaxies.
Here we introduce a semi-analytic model to quantify how rotating
outflows impact the morphology of the Lyman-$\alpha$ emission line.   
The model is contrasted against Monte Carlo radiative transfer
simulations of outflowing gas with additional solid body rotation.
We explore a range of neutral Hydrogen optical depth of
$10^5\leq\tauh\leq10^7$, rotational velocity $0\leq \vrot/\kms \leq
100$ and outflow velocity $0\leq \vout/\kms\leq 50$.  
We find three important consequences of rotation.
First, it introduces a dependency with viewing angle; second it
produces a line broadening and third it increases the flux at the
line's center.
The semi-analytic model reproduces the radiative transfer results for
the line width and flux change at the line's center within a $7\%$
and $50\%$ precision for an optical depth of
$\tauh=10^5$, respectively, and within $2\%$ and $1\%$ for an optical
depth of $\tauh=10^7$.
The semi-analytic model is a convenient tool to introduce rotational
kinematics as a post-processing step of idealized Monte Carlo
simulations, providing and easy way to include rotation in models that
seek to interpret \lya spectra in systems where rotation is evident
through kinematic maps. 
\end{abstract}
